---
title: "Practical Machine Learning - Prediction Assignment"
author: "Shree"
date: "November 23, 2016"
output: html_document
---

# ------------------------------------------------------------------------

### Introduction

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible 
to collect a large amount of data about personal activity relatively 
inexpensively. These type of devices are part of the quantified self movement – 
a group of enthusiasts who take measurements about themselves regularly to 
improve their health, to find patterns in their behavior, or because they are 
tech geeks. One thing that people regularly do is quantify how much of a 
particular activity they do, but they rarely quantify how well they do it. 
In this project, your goal will be to use data from accelerometers on the belt, 
forearm, arm, and dumbell of 6 participants. They were asked to perform barbell 
lifts correctly and incorrectly in 5 different ways. More information is 
available from the website here: 
http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting 
Exercise Dataset) 

---
---


### Data Source

---
---


The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv


The data for this project come from this source: 
http://groupware.les.inf.puc-rio.br/har. If you use the document you create for 
this class for any purpose please cite them as they have been very generous in 
allowing their data to be used for this kind of assignment.

---
---

### Loading the Datasete

---
---

First download the data files from the Internet and load them into two data 
frames. We will end up with a training dataset and a observations testing dataset.



```{r}
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(randomForest)
library(gbm)
library(plyr)
library(e1071)

# Downloading the training data

download.file(url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", 
              destfile = "c:/Users/imshr/Downloads/pml-training.csv")

# Load training dataset
training_dataSet <- read.csv("./pml-training.csv", na.strings=c("NA","#DIV/0!",""))

# Downloading the testing data
download.file(url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", 
              destfile = "c:/Users/imshr/Downloads/pml-testing.csv")

# Load testing dataset
testing_dataSet <- read.csv("./pml-testing.csv", na.strings=c("NA","#DIV/0!",""))



```

### Cleaning the Data

---
---

We will frist remove all columns that contains NA and remove features that are 
not in the testing dataset. The features containing NA are the variance, mean 
and standard devition (SD) within each window for each feature. Since the 
testing dataset has no time-dependence, these values are useless and can be 
disregarded. We will also remove the first 7 features since they are related to 
the time-series or are not numeric.

```{r}

features <- names(testing_dataSet[,colSums(is.na(testing_dataSet)) == 0])[8:59]

# Only use features used in testing cases.
training_dataSet <- training_dataSet[,c(features,"classe")]
testing_dataSet <- testing_dataSet[,c(features,"problem_id")]

dim(training_dataSet); dim(testing_dataSet);

```

---
---

### Partitioning the DataSet

---
---

Let's split the data into a training data set (60%) and a testing data set (40%). This will allow us to estimate the out of sample error of our predictor


```{r}

set.seed(12345)

inTrain <- createDataPartition(training_dataSet$classe, p=0.6, list=FALSE)
training <- training_dataSet[inTrain,]
testing <- training_dataSet[-inTrain,]

dim(training); dim(testing);

```


---
---

### Building the Decision Tree Model

---
---

Using Decision Tree, anything around 80% would be acceptable.

```{r}

set.seed(12345)
modFitDT <- rpart(classe ~ ., data = training, method="class", control = rpart.control(method = "cv", number = 10))
fancyRpartPlot(modFitDT)

```

---
---

### Predicting with the Decision Tree Model

---
---

```{r}

set.seed(12345)

prediction <- predict(modFitDT, testing, type = "class")
confusionMatrix(prediction, testing$classe)

```


---
---

### Building the Random Forest Model


Using random forest, the out of sample error should be small. The error will be estimated using the 40% testing sample. We should expect an error estimate of < 3%.


---
---

```{r}

set.seed(12345)

modFitRF <- randomForest(classe ~ ., data = training, method = "rf", importance = T, trControl = trainControl(method = "cv", classProbs=TRUE,savePredictions=TRUE,allowParallel=TRUE, number = 10))

plot(modFitRF)

```

---
---

### Building the Boosting Model

---
---

```{r}
modFitBoost <- train(classe ~ ., method = "gbm", data = training,
                    verbose = F,
                    trControl = trainControl(method = "cv", number = 10))

modFitBoost

plot(modFitBoost)

```


---
---

### Predicting with the Random Forest Model

---
---

```{r}

prediction <- predict(modFitRF, testing, type = "class")
confusionMatrix(prediction, testing$classe)

```


The random forest model performed very well in-sample, with about 99.3% Accuracy.


---
---

### Predicting with the Boosting Model


---
---

```{r}

prediction <- predict(modFitBoost, testing)
confusionMatrix(prediction, testing$classe)

```


---
---

## Predicting with the Testing Data 

---
---


### Decision Tree PRediction

---
---

```{r}

predictionDT <- predict(modFitDT, testing_dataSet)
predictionDT

```


---
---


### Random Forest PRediction

---
---

```{r}

predictionRF <- predict(modFitRF, testing_dataSet)
predictionRF

```

---
---


### Boosting PRediction

---
---

```{r}

predictionBoost <- predict(modFitBoost, testing_dataSet)
predictionBoost

```

---
---


## Submission 

---
---

As we can see from the confusion matrix the Random Forest model is very accurate, about 99%. Because of that we could expect nearly all of the submitted test cases to be correct. It turned out they were all correct.


```{r}


pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(predictionRF)


```








